---
title: "Grammer of Graphics"
subtitle: "Lecture 3"
date: "September 5th, 2022"
---

# Checklist {.smaller}

-- Have a GitHub account and have it connected with R (Lab 0)

-- Have checked the GitHub username sheet

-- Keeping up with the **Prepare** link on course website 

--   Go to the course [GitHub org](https://github.com/sta199-f22-2) and find your `ae-02-starwars` (repo name will be suffixed with your GitHub name).
--   Clone the repo in your container, open the Quarto document in the repo

## **Examining data visualization** {.smaller}

```{r}
#| echo: false
library(countdown)
```

::: columns
::: {.column width="50%"}
Discuss the following for the visualization.

--   What is the visualization trying to show?

--   What is effective, i.e. what is done well?

--   What is ineffective, i.e. what could be improved?

--   What are you curious about after looking at the visualization?

```{r}
#| echo: false
countdown(minutes = 4, left = 0, color_text = "#325b74")
```
:::


::: {.column width="50%"}

![Source: [Twitter](https://twitter.com/reina_sabah/status/1291509085855260672)](images/03-Grammer-of-Graphics/female-heights.png){fig-align="center"}

:::
:::


## Prepare Reading Questions

https://app.sli.do/event/g6fnKCDuZ1sw8NKiZTu3Rb

# Application exercise

## `ae-02-starwars`


-   Go to the course [GitHub org](https://github.com/sta199-f22-2) and find your `ae-02-starwars` (repo name will be suffixed with your GitHub name).
-   Clone the repo in your container, open the Quarto document in the repo, and follow along and complete the exercises.
-   Render, commit, and push your edits by the AE deadline -- 3 days from today (Thursday by 11:59p ET.


## Match the variables to plots

::: columns
::: {.column width="60%"}


--   1 categorical variable, 1 quantitative variable (often ordinal)

--   2 quantitative variables

--   1 quantitative variable 

```{r}
#| echo: false
countdown(minutes = 4, left = 0, color_text = "#325b74")
```
:::


::: {.column width="40%"}

-- geom_hist

-- geom_point

-- geom_bar

:::
:::






## Recap of AE {.smaller}

--   Construct plots with `ggplot()`.
--   Layers of ggplots are separated by `+`s.
--   The formula is (almost) always as follows:

```{r}
#| eval: false
ggplot(DATA, aes(x = X-VAR, y = Y-VAR, ...)) +
  geom_XXX()
```

--   Aesthetic attributes of a geometries (color, size, transparency, etc.) can be **map**ped to variables in the data or **set** by the user.

--   Use `facet_wrap()` when faceting (creating small multiples) by one variable and `facet_grid()` when faceting by two variables.
