---
title: "Types and Classes"
subtitle: "Lecture 8"
date: "September 21, 2022"
format: revealjs
---

## Checklist

```{r}
#| echo: false

library(countdown)
```

## Checklist 

-   Open your `ae-07` project in RStudio.

-   HW2 Due Tomorrow.

-   Exam 1 is coming (next week!)

## Announcements 

-- Regrade Requests

## Goals:

-- Understand how R treats data

-- ggplot practice

-- Practice re-creating graphs

## Warm Up: Types of variables

```{r}
#| echo: false

countdown(minutes = 3)
```


Categorical or quantitative?

-- Zipcode 

-- Ounces of water drank 

-- Height of students (measured tall / short)

-- Hours spent playing video games

# Application exercise

## `ae-07`

-   Go to the course [GitHub org](https://github.com/sta199-f22-2) and find your `ae-07` (repo name will be suffixed with your GitHub name).
-   Clone the repo in your container, open the Quarto document in the repo, and follow along and complete the exercises.
-   Render, commit, and push your edits by the AE deadline -- 3 days from today.

## Recap of AE {.smaller}

-- logical: a logical value.

-- integer: an integer (positive or negative). Many R programmers do not use this mode since every integer value can be represented as a double.

-- double: a real number stored in “double-precision floatint point format.”

-- complex: a complex number

-- character: a sequence of characters, called a “string” in other programming languages

-- list: a list of named values

-- NULL: a special type with only one possible value, known as NULL

More information can be found here: https://statsandr.com/blog/data-types-in-r/
