---
title: "AE 03: Visualizing penguins"
categories: 
  - Application exercise
editor: visual
execute:
  fig-width: 7
  fig-asp: 4
  fig-align: center
  warning: false
---

::: callout-important
Go to the [course GitHub organization](https://github.com/sta199-f22-2) and locate the repo titled `ae-03-data-types-YOUR_GITHUB_USERNAME` to get started.

This AE is due Sunday, Sep 10 at 11:59pm.
:::

For all analyses, we'll use the **tidyverse** and **palmerpenguins** packages.

```{r}
#| label: load-packages
#| warning: false

library(tidyverse)
library(palmerpenguins) #The data set name is penguins
```

The dataset we will visualize is called `penguins`. Let's `glimpse()` at it.

```{r}
#| label: glimpse

glimpse(penguins)
```

# Visualizing penguin weights - Demo

## Single variable

::: callout-note
Analyzing the a single variable is called **univariate** analysis.
:::

Create visualizations of the distribution of `weight`s of penguins.

1.  Make a histogram. Set an appropriate binwidth.

```{r}
#| label: histogram-weight

ggplot(penguins, 
       aes(x = body_mass_g)) +
       geom_histogram(binwidth = 500)

```

2.  Make a boxplot.

```{r}
#| label: boxplot-weight

ggplot(penguins, 
       aes(x = body_mass_g)) + 
  geom_boxplot()
```

3.  Based on these, determine if each of the following statements about the shape of the distribution is true or false.
    -   The distribution of penguin weights in this sample is left skewed.
    -   The distribution of penguin weights in this sample is unimodal (roughly symmetric).

## Two variables

::: callout-note
Analyzing the relationship between two variables is called **bivariate** analysis.
:::

Create visualizations of the distribution of `weight`s of penguins by `species`. Note: `aes`thetic is a visual property of one of the objects in your plot. Aesthetic options are:

-   shape
-   color
-   size
-   fill

4.  Make a single histogram. Set an appropriate binwidth.

```{r}
#| label: histogram-weight-species
ggplot(penguins, 
       aes(x = body_mass_g, fill = species )) +
       geom_histogram(binwidth = 250, alpha = .5)

```

5.  Use multiple histograms via faceting, one for each species. Set an appropriate binwidth, add color as you see fit, and turn off legends if not needed.

```{r}
#| label: histogram-weight-species-facet

ggplot(penguins, 
       aes(x = body_mass_g, fill = species )) +
       geom_histogram(binwidth = 250, alpha = .5, show.legend = FALSE) + 
      facet_wrap("species", ncol = 1)
```

6.  Use side-by-side box plots. Add color as you see fit and turn off legends if not needed.

```{r}
#| label: boxplot-weight-species

ggplot(penguins, 
       aes(x = species, y = body_mass_g,  fill = species)) + 
       geom_boxplot(show.legend = F)
```

7.  Use density plots. Add color as you see fit.

```{r}
#| label: density-weight-species

ggplot(penguins, 
       aes(x = body_mass_g,  fill = species)) +
       geom_density(alpha = .5)
```

8.  Use violin plots. Add color as you see fit and turn off legends if not needed.

```{r}
#| label: violin-weight-species

ggplot(penguins, 
       aes(x = species, y = body_mass_g,  fill = species)) + 
  geom_violin(alpha = 0.5 , show.legend = F)
```

9.  Make a jittered scatter plot. Add color as you see fit and turn off legends if not needed.

```{r}
#| label: jitter-weight-species


ggplot(penguins, 
       aes(x = species, y = body_mass_g, color = species)) +
  geom_jitter(show.legend = FALSE)
```

10. Use beeswarm plots. Add color as you see fit and turn off legends if not needed.

```{r}
#| label: beeswarm-weight-species

library(ggbeeswarm)

ggplot(penguins, 
       aes(x = species, y = body_mass_g, color = species)) +
  geom_beeswarm(show.legend = FALSE)

```

11. Use multiple geoms on a single plot. Be deliberate about the order of plotting. Change the theme and the color scale of the plot. Finally, add informative labels. Hint: scale_color_viridis: https://ggplot2.tidyverse.org/reference/scale_viridis.html theme: https://ggplot2.tidyverse.org/reference/ggtheme.html

```{r}
#| label: multi-geoms
ggplot(data = penguins, 
       mapping = aes(x = body_mass_g, y = species, color = species)) +
  geom_boxplot(binwidth = 500) +
  geom_jitter() +
  scale_color_viridis_d(option = "D", end = 0.8) +
  theme_minimal() +
  labs(x= "Weight",
       y = "Species", 
       title= "Weight Disrtribution of Penguins") +
  theme(legend.position = "None")

```

## Multiple variables

::: callout-note
Analyzing the relationship between three or more variables is called **multivariate** analysis.
:::

12. Facet the plot you created in the previous exercise by `island`. Adjust labels accordingly.

```{r}
#| label: multi-geoms-facet
ggplot(data = penguins, 
       mapping = aes(x = body_mass_g, y = species, color = species)) +
  geom_boxplot(binwidth = 500) +
  geom_jitter() +
  scale_color_viridis_d(option = "D", end = 0.8) +
  theme_minimal() +
  labs(x= "Weight(g)",
       y = "Species", 
       title= "Weight Disrtribution of Penguins") +
  theme(legend.position = "None") +
  facet_wrap("island")
```

Before you continue, let's turn off all warnings the code chunks generate and resize all figures. We'll do this by editing the YAML.

## Visualizing other variables - Your turn! Optional

13. Pick a single categorical variable from the data set and make a bar plot of its distribution.

```{r}
#| label: barplot
ggplot(penguins, 
       aes(x = species)) +
  geom_bar(show.legend = T) +
  scale_color_viridis_d(option = "D", end = 0.8) +
  theme_minimal() +
  labs(
    x = "Species by Sex",
    title = "Penguins by species"
  )

```

14. Pick two categorical variables and make a visualization to visualize the relationship between the two variables. Along with your code and output, provide an interpretation of the visualization.

```{r}
#| label: cat-cat

ggplot(penguins, 
       aes(x = species, fill = sex)) +
  geom_bar(show.legend = T) +
  scale_color_viridis_d(option = "D", end = 0.8) +
  theme_minimal() +
  labs(
    x = "Species by Sex",
    title = "Penguins by species and sex"
  )
```

It appears that there is little relationship between the type of species and sex of species. Across all three species, the sex for each species seems to be fairly proportional.

15. Make another plot that uses at least three variables. At least one should be numeric and at least one categorical. In 1-2 sentences, describe what the plot shows about the relationships between the variables you plotted. Don't forget to label your code chunk.

```{r}
ggplot(penguins, 
       aes(x = species, y = body_mass_g, color = species)) +
  geom_beeswarm(show.legend = FALSE) +
  geom_boxplot(show.legend = FALSE, alpha = 0.5) +
  facet_wrap(~island) +
  scale_color_viridis_d(option = "D", end = 0.8) +
  theme_minimal() +
  labs(
    x = "Species",
    y = "Body mass (g)",
    title = "Body mass of penguins in Palmer Archipelago",
    subtitle = "By species and island"
  )
```

It appears that the body mass of penguins depends on both the species of the penguin and the island the penguin is on. The Gentoo penguins on the Biscoe island are much larger than the others.



