---
title: "AE 01: MTCars"
categories: 
  - Application exercise
editor: visual
execute:
  fig-width: 7
  fig-asp: 4
  fig-align: center
  warning: false
---


These data were extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973--74 models).

To learn more about the data, see here: https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/mtcars

```{r}
library(tidyverse)
data("mtcars")
```

1.  Create a code chunk and run the *summary* command on the variable mpg.

```{r}

summary(mtcars$mpg)

```

2.  Fill in the following code (replace the ... below) to create a histogram of mpg from question 1.

```{r}
ggplot(mtcars, aes(x= mpg)) + geom_histogram()
```

Did it work? If not, you may need to library the ggplot2 package. do that in the code above!

3.  That histogram is really boring.... fill in your same x and give it an informative title by filling the " " below.

```{r}
ggplot(mtcars, aes(x= mpg)) + geom_histogram() + labs(title="Histogram of mpg")
```

4.  Let's explore if cars with or without a transmission get better mpg. Put am in the space for color and fill and run the code. Continue to put your variable in x = and write out a title.

```{r}
mtcars$am <- as.factor(mtcars$am) #This makes sure am is a factor

ggplot(mtcars, aes(x= mpg, color = am , fill = am )) + geom_histogram(binwidth = 1) + labs(title="Histogram of mpg by transmission")

```

5.  Let's add a density curve. Run the following code below. Play around with alpha and see how the graph changes.

```{r}
ggplot(mtcars, aes(x= mpg , color = am , fill = am )) + geom_histogram(aes(y=..density..), position="identity") + labs(title="MPG") +
geom_density(alpha=0.6)
```

6.  Let's rip the histogram apart and seperate by the variable am. Fill in your quantitative variable in x = and put your transmission variable in the ... after facet_grid.

```{r}
ggplot(mtcars, aes(x= mpg))+
  geom_histogram(color="black", fill="white")+
  facet_grid(~am)
```

7.  For additional practice with the MT data set, please click [here](https://www.statology.org/mtcars-dataset-r/)
