---
title: "Modeling Introduction"
format: html
editor: visual
---

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
```

```{r}
#| message: false
fish <- read_csv("data/Fish.csv")
```

For this activity, we will be working with a Fish data set. This data set is a record of 2 common different fish species in fish market sales. Documentation of the variables can be seen below:

| **variable** | **description**         |
|--------------|-------------------------|
| `Species`    | species name of fish    |
| `Weight`     | weight of fish in grams |
| `Length1`    | vertical length in cm   |
| `Length2`    | diagonal length in cm   |
| `Length3`    | cross length in cm      |
| `Height`     | height in cm            |
| `Width`      | diagonal width in cm    |

We are going to investigate the relationship between a fish's height and weight. Below, create an appropriate plot to investigate this relationship. Add appropriate labels to the plot.

```{r}
fish |> 
  ggplot(
    aes(x = Height, y = Weight)
  ) + 
  geom_point() +
  labs( 
    title = "Fish Height by Weight")
```

-   *(2-min)* If you were to draw a a straight line to best represent the relationship between a fish's height and width, how would you do it?

Start from the bottom and go up Identify the first and last point and draw a line through most the others

-   Now, let R draw the line for us. Please refer to the documentation here: https://ggplot2.tidyverse.org/reference/geom_smooth.html. Specifically, refer to the `method` section.

```{r}
fish |> 
  ggplot(
    aes(x = Height, y = Weight)
  ) + 
  geom_point() +
  geom_smooth(method = "lm") +
  labs( 
    title = "Fish Height by Weight")
```

-   What types of questions can this plot help answer?

Is there a relationship between fish height and weight?

Predict

-   *(2-min)* We can use this line to make predictions. Predict what you think the width of a fish would be with a height of 10 cm. 15cm. 20cm. Which prediction is considered extrapolation?

At 10cm, we estimate a weight of 375 grams At 15cm, we estimate a weight of 600 grams At 20cm, we estimate a weight of 975 grams

-   Brainstorm. How was this line created?

R probably created a line that best fit the data....but how and what does that mean?

-   What is a residual?

The difference between what we observe and what our model predicts

-   **Demo:**

```{r}
#Modeling Code! More on this Monday!

fit <- lm(Width ~ Height, data = fish)
fish$predicted <- predict(fit)
fish$residuals <- residuals(fit)


fish |> 
ggplot(aes(x = Height, y = Width)) + 
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +  
  geom_segment(aes(xend = Height, yend = predicted), alpha = .2) +  
  geom_point() + 
  geom_point(aes(y = predicted), shape = 1) + 
  theme_bw()  
```

Now we know how the line is fit.... so what's the line?

```{r}
#| eval: false
fish |>
  ggplot(
    aes(x = Height, y = Weight)
  ) + 
  geom_point() + 
  geom_smooth(method = "lm", fullrange = T) +
  labs(
    title = "Fish by Height and Width"
  ) + 
  xlim(0,20)

summary(fit)
```

-   *(2-min)* Using the line, predict what you think the width of a fish would be with a height of 10 cm. 15cm. 20cm.

$\hat{weight}$ = 1.96 + 0.23*10
$\hat{weight}$ = 1.96 + 0.23*15
$\hat{weight}$ = 1.96 + 0.23*20

-   What is the notation of our line?

$$

\hat{weight} = 1.96 +0.23*height

$$


## Correlation

We can also assess correlation between two quantitative variables.

What is correlation? What is correlation bounded by?

*Strength and direction of a linear relationship*

r = \[-1, 1\]

What is the correlation between Height and Weight?

```{r}
fish |>
  select(Height, Weight) |>
  cor()
```


Are you good at guessing? Give it a try!

https://www.rossmanchance.com/applets/2021/guesscorrelation/GuessCorrelation.html

## Modeling a third variable

Does the relationship change by species? Plot two separate straight lines by Bream and Roach species of fish below.

```{r}
fish |> 
  ggplot(
    aes(x = Height, y = Weight, color = Species)
  ) + 
  geom_point() +
  geom_smooth(method = "lm") +
  labs( 
    title = "Fish Height by Weight")
```

-   We can fit more models than just a straight line. Change the following code below to read `method = loess`. What is different from the plot created before?

The line isn't straight!

```{r}
#| eval: false
fish |>
  ggplot(
    aes(x = Height, y = Width)
  ) + 
  geom_point() + 
  geom_smooth(method = "loess") +
  labs(
    title = "Fish by Height and Width"
  )
```
