---
title: "Inference overview + Review"
format: html
editor: visual
---

```{r}
#| label: packages
#| message: false

library(tidyverse)
library(tidymodels)
```

```{r}
#| label: data
#| message: false

rc <- read_csv("data/roller_coasters.csv")
```

In this application exercise, we will be looking at a roller coaster and amusement park database by Duane Marden.[^1] This database records multiple features of roller coasters. For the purpose of this activity, we will work with a random sample of 157 roller coasters.

[^1]: [Roller Coaster DataBase](https://rcdb.com/), [UltimateRollerCoaster.com](https://www.ultimaterollercoaster.com/), [ESTEEM: Enhancing Statistics Teachers Education with E-Modules](https://www.fi.ncsu.edu/teams/hirise/).

+---------------------+--------------------------------------------------------------------------+
| **Variable**        | **Description**                                                          |
+:====================+:=========================================================================+
| `age_group`         | 1: Older (Built between 1900-1979)                                       |
|                     |                                                                          |
|                     | 2: Recent (1980-1999)                                                    |
|                     |                                                                          |
|                     | 3: Newest (2000-current)                                                 |
+---------------------+--------------------------------------------------------------------------+
| `coaster`           | Name of the roller coaster                                               |
+---------------------+--------------------------------------------------------------------------+
| `park`              | Name of the park where the roller coaster is located                     |
+---------------------+--------------------------------------------------------------------------+
| `city`              | City where the roller coaster is located                                 |
+---------------------+--------------------------------------------------------------------------+
| `state`             | State where the roller coaster is located                                |
+---------------------+--------------------------------------------------------------------------+
| `type`              | Material of track (Steel or Wooden)                                      |
+---------------------+--------------------------------------------------------------------------+
| `year_opened`       | Year when roller coaster opened                                          |
+---------------------+--------------------------------------------------------------------------+
| `top_speed`         | Maximum speed of roller coaster (mph)                                    |
+---------------------+--------------------------------------------------------------------------+
| `max_height`        | Highest point of roller coaster (ft)                                     |
+---------------------+--------------------------------------------------------------------------+
| `drop`              | Length of largest gap between high and low points of roller coaster (ft) |
+---------------------+--------------------------------------------------------------------------+
| `length`            | Length of roller coaster track (ft)                                      |
+---------------------+--------------------------------------------------------------------------+
| `duration`          | Time length of roller coaster ride (sec)                                 |
+---------------------+--------------------------------------------------------------------------+
| `inversions`        | Whether or not roller coaster flips passengers at any point (Yes or No)  |
+---------------------+--------------------------------------------------------------------------+
| `num_of_inversions` | Number of times roller coaster flips passengers                          |
+---------------------+--------------------------------------------------------------------------+

## Practice with data manipulation

-   **Demo:** Create a new variable called `opened_recently` that has a response of `yes` if the roller coaster was opened after 1970, and `no` if the roller coaster was built during or before 1970. Save this new factor variable in the `rc` data set. Additionally, make `type` a factor variable.

```{r}
#| label: data-prep

rc <- rc |>
  mutate(
    opened_recently = if_else(year_opened > 1970, "yes", "no"), 
    opened_recently = as.factor(opened_recently),
    type = as.factor(type)
  ) 
```

> **We want to investigate the relationship between how fast a roller coaster goes, and how long a roller coaster lasts. Specifically, we are interested in how well the duration of a roller coaster explains how fast it is.**

-   **Your turn:** Based on this research question, which two variables should we use from our data set? Which is our response?

top_speed; duration

top_speed is our response

-   **Your turn:** Describe how we can construct a bootstrap distribution for the slope of the model predicting `top_speed` from `duration`.

-   **Demo:** Now, construct this bootstrap distribution.

```{r}
#| label: boot-df-slope

set.seed(12345)

boot_df_slope <- rc |>
  specify(top_speed ~ duration) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "slope")
```

-   **Demo:** Create a 90% confidence interval by filling in the code below.

```{r}
#| label: slope-ci

boot_df_slope |>
  summarize(
    lower = quantile(stat, 0.05),
    upper = quantile(stat, 0.95)
  )
```

-   **Demo:** Interpret the confidence interval in the context of the problem.

We are 90% confident that for a 1 min increase in duration, the true/population mean speed (y) is expected to be 0.0195 lower to 0.132 mph higher.

-   **Demo:** Prediction: How is this different than what we've done before?

Now let's practice predicting!

-   **Demo:** Report the tidy output of your model below

```{r}
#| label: speef-fit

speed_fit <- linear_reg() |>
  set_engine("lm") |>
  fit(top_speed ~ duration, data = rc) 

tidy(speed_fit)
```

-   Below is the estimated model in proper notation.

$\widehat{top~speed} = 52 + 0.056 \times duration$

-   **Demo:** Use your model to estimate the top speed of a roller coaster if their duration is 155 minutes

```{r}
#| label: prediction

duration_155 <- tibble(duration = 155)

predict(speed_fit, new_data = duration_155)
```

## Interpret

-   Now, interpret the slope coefficient.

For a 1 unit minute increase in duration, we estimate on average a 0.0560 increase in mph.

-   What if our response was logged? How does the interpretation change?

$\widehat{log(TopSpeed)} = 52 + 0.056 \times Duration$

Recall that to interpret the slope value we need to exponentiate it if we want to be back on the original scale!

Why?

$log(y) = \beta_o + \beta_1x$

y = $exp(\beta_o + \beta_1x)$)

y = $exp(\beta_o) * exp(\beta_1x)$)

This implies that our explanatory variable has a multiplicative relationship with our response variable

*For a one minute increase in duration, we estimate on average, top speed to increase by a factor of 0.05, or by about 5%.*

# Research question 3

> **We are also interested in investigating if roller coasters after 1970 are faster than those opened before. For this question, we want to estimate the difference.**

-   Should we make a confidence interval or hypothesis test?

-   **Demo:** Now, use bootstrapping to estimate the difference between the true speeds of roller coasters before and after 1970.

```{r boot-sample}
#| label: two-mean-boot

set.seed(12345)

boot_df <- rc |>
  specify(top_speed ~ opened_recently) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "diff in means", order = c("yes", "no"))
```

-   **Your turn:** Create a 99% confidence interval and interpret it in the context of the data and the research question.

```{r}
#| label: two-mean-ci

boot_df |>
  summarize(
    lower = quantile(stat, 0.005),
    upper = quantile(stat, 0.995)
  )
```

ORDER MATTERS (yes - no)

We are 99% confident that the true mean speed for roller coasters opened after 1970 ($\mu_1$) is (9.15, 23) mph LARGER than the true mean speed for roller coasters opened before 1970 ($\mu_2$)

# Discussion

-   Brainstorm the different between bootstrap and permute. Write down the key differences below. *Hint:* Think back to the different sampling techniques between question 1 and question 2 / 3.

The main difference between permutation and bootstrapping is that bootstrapping resamples, with replacement from our data. Permutation can be thought of as shuffling / distributing our data / resampling without replacement.

## Interpretation - Return of the Penguins

We will fit a SRL, MLR, and logistic regression model and compare how we interpret the slope coefficient.

```{r}
library(palmerpenguins)
```

```{r}
glimpse(penguins)
```

### Simple Linear Regression

```{r}
modelSLR <- linear_reg() |>
  set_engine("lm") |>
  fit(flipper_length_mm ~ bill_length_mm, data = penguins)

tidy(modelSLR)
```

-   Interpret the slope coefficient assocciated with bill_length_mm in the context of the problem

For a 1 mm increase in bill length, we estimate on average, flipper length to increase by 1.69 mm.

*Add response*

### Multiple Linear Regression

```{r}
modelMLR <- linear_reg() |>
  set_engine("lm") |>
  fit(flipper_length_mm ~ bill_length_mm + island, data = penguins)

tidy(modelMLR)
```

-   Interpret the slope coefficient associated with bill_length_mm in the context of the problem

For a 1 mm increase in bill length, we estimate on average, flipper length to increase by 1.51 mm, holding island constant.

### Logistic Regression

```{r}

modellog <- logistic_reg() |>
  set_engine("glm") |>
  fit(sex ~ bill_length_mm, data = penguins)

tidy(modellog)
```

-   Interpret the slope coefficient associated with bill_length_mm in the context of the problem and on the original scale

For a 1 unit increase in bill length, we expect the odds of being female to increase by a factor of .14, or by about 14%.
